---
title: "scnanoseq - 10X 3' and 5': Figures for manuscript"
author:
  - name: "Austyn Trull"
  - name: "Lara Ianov, Ph.D."
output:
  html_document:
    code_folding: hide
    css: "style.css"
    toc: true
    toc_float: true
---

Task: final figures for the manuscript (previous figs were for initial analysis. In this script, a subset is refined).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
set.seed(1234)

library(ggplot2)
library(Seurat)
library(dplyr)
library(grid)
library(tidyr)

# add custom functions from UBDS R scripts
lapply(list.files("./scripts/R"), FUN = function(x) source(paste0("./scripts/R/", x)))

#### CONSTANT DEFS ####

## The prefix to output files
output_prefix <- "./output_10X_3_5_prime_dataset_figs"

dir.create(file.path(output_prefix), recursive = TRUE)
```

# Load pre-processed Seurat objects

```{r}
seurat_int <- readRDS(file = "./output_scnanoseq_10X_data/seurat_int_gene_level.rds")
seurat_objs_transcript <- readRDS(file = "./output_scnanoseq_10X_data/seurat_objs_transcript_level.rds")

# Set default Idents

cell_idents <- c("predicted.celltype.l2", "predicted.ann_level_3")

seurat_int <- mapply(FUN = function(x,y) {

  Idents(x) <- y
  
  return(x)
  
}, x=seurat_int, y=cell_idents)

seurat_objs_transcript <- mapply(FUN = function(x,y) {

  Idents(x) <- y
  
  return(x)
  
}, x=seurat_objs_transcript, y=cell_idents)

# Add sample metadata to integrated object which is more user-friendly for readers

seurat_int$SC3pv3_gene_int$sample_info <- ifelse(seurat_int$SC3pv3_gene_int@meta.data$orig.ident == "SC3pv3_gene_isoquant", yes = "scnanoseq_Oxford", no = "CR_Illumina")
seurat_int$SC5pv2_gene_int$sample_info <- ifelse(seurat_int$SC5pv2_gene_int@meta.data$orig.ident == "SC5pv2_gene_isoquant", yes = "scnanoseq_Oxford", no = "CR_Illumina")
```

# UMAPs (Gene-level)

```{r}
mapply(FUN = function(x, y) {
  
  if (y == "SC3pv3_gene_int") {
    cell_prediction <- "predicted.celltype.l2"
    title <- paste0("Gene-level 3' PBMC")
    out_file <- paste0(output_prefix, "/UMAP_gene_level_PBMC.pdf")
  } else {
    cell_prediction <- "predicted.ann_level_3"
    title <- paste0("Gene-level 5' lung cancer DTCs")
    out_file <- paste0(output_prefix, "/UMAP_gene_level_lung.pdf")
  }
  
  p1 <- DimPlot(x,
                group.by = cell_prediction,
                split.by = "sample_info",
                reduction = "umap",
                label = TRUE,
                pt.size = NULL,
                label.size = 4,
                repel = TRUE) + NoLegend() +
    guides(color = guide_legend(override.aes = list(size = 4), ncol=1)) +
    ggtitle(title) +
    theme(axis.title = element_text(size = 20),
          axis.text.x = element_text(size=15),
          axis.text.y = element_text(size=15),
          plot.title = element_text(size = 30, color = "#0072B2"),
          strip.text.x = element_text(size = 20, face = "bold"))
  #save
  
  pdf(out_file, height = 8, width = 8)
  plot(p1)
  dev.off()
  
}, x = seurat_int, y = names(seurat_int))
```

# UMAPs (Transcript-level)

```{r}
mapply(FUN = function(x, y) {
  
  if (y == "SC3pv3_transcript_isoquant" || y == "SC3pv3_transcript_oarfish") {
    cell_prediction <- "predicted.celltype.l2"
    
    if (y == "SC3pv3_transcript_isoquant") {
      out_file <- paste0(output_prefix, "/UMAP_transcript_level_PBMC_isoquant.pdf")
      title <- paste0("Transcript-level 3' PBMC IsoQuant")
    } else {
      out_file <- paste0(output_prefix, "/UMAP_transcript_level_PBMC_oarfish.pdf")
      title <- paste0("Transcript-level 3' PBMC oarfish")
    }
    
  } else {
    cell_prediction <- "predicted.ann_level_3"
    
    if (y == "SC5pv2_transcript_isoquant") {
      out_file <- paste0(output_prefix, "/UMAP_transcript_level_lung_isoquant.pdf")
      title <- paste0("Transcript-level 3' PBMC IsoQuant")
    } else {
      out_file <- paste0(output_prefix, "/UMAP_transcript_level_lung_oarfish.pdf")
      title <- paste0("Transcript-level 3' PBMC oarfish")
    }
    
  }
  
  p1 <- DimPlot(x,
                group.by = cell_prediction,
                #split.by = "sample_info",
                reduction = "umap",
                label = TRUE,
                pt.size = NULL,
                label.size = 4,
                repel = TRUE) + NoLegend() +
    guides(color = guide_legend(override.aes = list(size = 4), ncol=1)) +
    ggtitle(title, subtitle = "scnanoseq_Oxford") + 
    theme(axis.title = element_text(size = 20),
          axis.text.x = element_text(size=15),
          axis.text.y = element_text(size=15),
          plot.title = element_text(size = 30, color = "#009E73"),
          plot.subtitle = element_text(size = 20, face = "bold", hjust = 0.5))
  #save
  
  pdf(out_file, height = 8, width = 8)
  plot(p1)
  dev.off()
  
}, x = seurat_objs_transcript, y = names(seurat_objs_transcript))
```

# nFeature Correlation plot (Gene-level, pre-filtering)

Same plot generated during the analysis, with minor cosmetic changes below for paper:

```{r}
# import data (pre-filtering)

pre_filter_objs <- readRDS("./output_scnanoseq_10X_data/all_10X_prefilter.rds")
```

```{r}
# merge metadata based on barcodes from scnanoseq and CR runs (gene-level only)

merged_meta <- mapply(FUN = function(x,y) {
  
  scnano_meta <- x@meta.data
  CR_meta <- y@meta.data
  
  # add BCs as column
  scnano_meta$cell_barcodes <- rownames(scnano_meta)
  CR_meta$cell_barcodes <- rownames(CR_meta)
  
  # 10X CR contains -1 (data imported by keeping them for now, but need to remove for this step)
  
  CR_meta$cell_barcodes <- gsub("-1","",CR_meta$cell_barcodes)
  
  # merge by keeping all barcodes (not just what is common b/t the two)
  merged_meta <- merge(x = scnano_meta,
                       y = CR_meta,
                       by = "cell_barcodes",
                       suffixes= c("_scnanoseq", "_ground_truth"),
                       all = TRUE)
  
  # Change NA to 0 to display any barcodes unique to each dataset
  merged_meta <- merged_meta %>% mutate_if(is.numeric, ~replace_na(., 0))
  
}, x=pre_filter_objs[c(1,3)], y=pre_filter_objs[c(7,8)], SIMPLIFY = FALSE)
```

```{r}
## nFeature scatters ##

# params

feature_1 <- "nFeature_RNA_scnanoseq"
feature_2 <- "nFeature_RNA_ground_truth"
output_feature <- "_nFeature_pre_filter.pdf"

mapply(FUN = function(x,y) {
  
  scatter_plot <- scatter_corr(data.frame = x,
                               feature = feature_1,
                               feature2 = feature_2,
                               feature_type = "gene",
                               title_size = 30,
                               axis_size = 15,
                               point_size = 2,
                               point_alpha = 0.5) + 
    theme(axis.title = element_text(size = 20))
  
  # make axis titles cleaner:
  
  scatter_plot <- scatter_plot + xlab("nFeature scnanoseq_Oxford") + ylab("nFeature CR_Illumina")
  
  file_name <- paste0(output_prefix, "/", y, output_feature)

  #adding some room to the right to avoid axis being cutoff
  pdf(file = file_name, height = 8, width = 8)
  plot(scatter_plot + theme(plot.margin = margin(t = 20, b = 20, l = 50, r = 50)))
  dev.off()
  
}, x=merged_meta, y=names(merged_meta))
```

# Marker genes (Gene-level)

```{r}
marker_pbmc <- read.csv("./input/references/markers_subset_pbmc.csv")
marker_lung <- read.csv("./input/references/markers_subset_lung.csv")
```

```{r}
mapply(FUN = function(x, y) {
  
  if (y == "SC3pv3_gene_int") {
    cell_prediction <- "predicted.celltype.l2"
    title <- paste0("Gene-level 3' PBMC")
    out_file <- paste0(output_prefix, "/Violins_gene_level_PBMC.pdf")
    marker_input <- marker_pbmc$markers
  } else {
    cell_prediction <- "predicted.ann_level_3"
    title <- paste0("Gene-level 5' lung cancer DTCs")
    out_file <- paste0(output_prefix, "/Violins_gene_level_lung.pdf")
    marker_input <- marker_lung$markers
  }
  
  p1 <- VlnPlot(object = x,
                split.by = "sample_info",
                group.by = cell_prediction,
                features = marker_input,
                assay = "SCT",
                stack = TRUE,
                flip = TRUE) + 
    ggtitle(title) +
    theme(legend.position="bottom",
          legend.text = element_text(size = 15),
          axis.title = element_text(size = 20),
          axis.text.x = element_text(size=15),
          axis.text.y = element_text(size=15))
  
  #save
  
  pdf(out_file, height = 10, width = 10)
  plot(p1)
  dev.off()
  
}, x = seurat_int, y = names(seurat_int))
```

# Marker genes (Transcript-level)

```{r}
# search for all transcript under a gene name withing each quantifier
transcript_list <- mapply(FUN = function(x,y) {
  
  if (y == "SC3pv3_transcript_isoquant" || y == "SC3pv3_transcript_oarfish") {
    
    markers <- marker_pbmc$markers
    
  } else {
    markers <- marker_lung$markers
    
  }
  
  transcripts <- list()
  
  # Searches for transcript features within SCT counts
  for (marker in markers) {
    feature <- sort(grep(pattern = paste0("^", marker, "\\."),
                             x = rownames(x@assays$SCT@counts),
                             value = TRUE))
    
    transcripts <- c(transcripts, feature)
    
  }
  
  transcripts <- transcripts[lengths(transcripts) > 0]
  
  transcripts <- as.character(unlist(transcripts))
  
  return(transcripts)
  
}, x = seurat_objs_transcript, y = names(seurat_objs_transcript))

# remove any transcripts where sum == 0

transcript_list <- mapply(FUN = function(x,y) {
  
  x <- x[sapply(x, function(transcript) {
    sum(GetAssayData(y, layer = "data")[transcript, ]) > 0
  })]
  
}, x = transcript_list, y = seurat_objs_transcript)
```


### Plotting markers for cells with no labels ###
Given the starting point of this analysis with scnanoseq outputs (isoquant outs [gene and transcript] and oarfish [transcript only at this time]), this validation analysis used
the gene level data to transfer labels to both transcript datasets. However, in the oarfish data, a small amount of cells do not have labels in cases where a cell was retained in oarfish but
not in isoquant's downstream analysis.

**NOTE**: For simplicity in the scope of this analysis, the dotplot figures from the section below removes the subset of cells which do not contain a label. However, in order to visualize that these cells
can still be labelled by marker expression, we plot this subset of cells here with their respective markers. 
Thus, for more in-depth tertiary analysis we recommend users to label all cells (in this case one can manually overwrite any cells without labels to contain the labels where the markers expression can define the cell type.

This is not done in this case, since we re-use these scripts to validate the pipeline with each release and the seurat cluster numbers can change from run to run.

```{r}
# see issue https://github.com/satijalab/seurat/issues/8772 where dotplot doesn't plot cells which do not contain a label,
# thus we separately visualize them here with seurat clusters instead. 
# Based on this any NA cells can be further annotated, but for the scope of this analysis, only this visualization is done

# In 3', 126 cells do not contain a label and 5' has 36
seurat_transcript_tmp <- mapply(FUN = function(x, y) {

  if (y == "SC3pv3_transcript_oarfish") {
    cell_prediction <- "predicted.celltype.l2"
  } else {
    cell_prediction <- "predicted.ann_level_3"
  }
  
  na_cells <- x@meta.data %>%
    dplyr::select(.data[[cell_prediction]]) %>%
    dplyr::filter(is.na(.data[[cell_prediction]]))
  
  message("---Number of cells without a label is:---")
  print(nrow(na_cells))
  
  obj <- subset(x, cells = rownames(na_cells))
  
}, x = seurat_objs_transcript[c(2,4)], y = names(seurat_objs_transcript[c(2,4)]))

# filter transcript further to remove any not present in na cells
transcript_list_subset <- mapply(FUN = function(x,y) {
  
  x <- x[sapply(x, function(transcript) {
    sum(GetAssayData(y, layer = "data")[transcript, ]) > 0
  })]
  
}, x = transcript_list[c(2,4)], y = seurat_transcript_tmp)

# plot
mapply(FUN = function(x, y) {
  
  if (y == "SC3pv3_transcript_oarfish") {
    out_file_dotplot <- paste0(output_prefix, "/PBMC_na_cells_dotplot.pdf")
    out_file_featureplot <- paste0(output_prefix, "/PBMC_na_cells_featureplot.pdf")
    marker_input <- transcript_list_subset$SC3pv3_transcript_oarfish
    width_out <- 14
  } else {
    out_file_dotplot <- paste0(output_prefix, "/lung_na_cells_dotplot.pdf")
    out_file_featureplot <- paste0(output_prefix, "/lung_na_cells_featureplot.pdf")
    marker_input <- transcript_list_subset$SC5pv2_transcript_oarfish
    width_out <- 10
  }
  
  Idents(x) <- "seurat_clusters"
  
  p1 <- DotPlot(object = x,
                group.by = "seurat_clusters",
                features = rev(marker_input),
                assay = "SCT",
                cols = c("#CC79A7","#009E73")) + 
    coord_flip() +
    theme(axis.title = element_text(size = 20),
          axis.text.x = element_text(size=15, angle = 45, hjust=1),
          axis.text.y = element_text(size=15))
  
  pdf(out_file_dotplot, height = 16, width = width_out)
  plot(p1)
  dev.off()
  
  # to co-visualize expression with cell labels
  p2 <- FeaturePlot(object = x,
                    features = rev(marker_input),
                    cols = c("#D3D3D3", "#009E73"),
                    order = TRUE)
  
  pdf(out_file_featureplot, width = 20, height = 20)
  plot(p2)
  dev.off()
  
}, x = seurat_transcript_tmp, y = names(seurat_transcript_tmp))

rm(seurat_transcript_tmp, transcript_list_subset)
```


### DotPlots (transcript-level)

```{r}
# given visual inspection of the above, we now plot all transcript while removing
# the few cells which do not have a label from oarfish analysis
seurat_transcript_tmp <- seurat_objs_transcript

seurat_transcript_tmp$SC3pv3_transcript_oarfish <- subset(seurat_transcript_tmp$SC3pv3_transcript_oarfish,
                                                          predicted.celltype.l2 %in% na.omit(seurat_transcript_tmp$SC3pv3_transcript_oarfish$predicted.celltype.l2))

seurat_transcript_tmp$SC5pv2_transcript_oarfish <- subset(seurat_transcript_tmp$SC5pv2_transcript_oarfish,
                                                          predicted.ann_level_3 %in% na.omit(seurat_transcript_tmp$SC5pv2_transcript_oarfish$predicted.ann_level_3))


mapply(FUN = function(x, y) {
  
  if (y == "SC3pv3_transcript_isoquant" || y == "SC3pv3_transcript_oarfish") {
    cell_prediction <- "predicted.celltype.l2"
    out_file <- paste0(output_prefix, "/DotPlot_PBMC_", y,".pdf")
    width_out <- 14
    
    if (y == "SC3pv3_transcript_isoquant") {
      marker_input <- transcript_list$SC3pv3_transcript_isoquant
      title <- paste0("Transcript-level 3' PBMC (IsoQuant)")
    } else {
      marker_input <- transcript_list$SC3pv3_transcript_oarfish
      title <- paste0("Transcript-level 3' PBMC (oarfish)")
    }
    
  } else {
    cell_prediction <- "predicted.ann_level_3"
    out_file <- paste0(output_prefix, "/DotPlot_lung_", y,".pdf")
    width_out <- 10
    
    if (y == "SC5pv2_transcript_isoquant") {
      marker_input <- transcript_list$SC5pv2_transcript_isoquant
      title <- paste0("Transcript-level 5' lung cancer DTCs (IsoQuant)")
    } else {
      marker_input <- transcript_list$SC5pv2_transcript_oarfish
      title <- paste0("Transcript-level 5' lung cancer DTCs (oarfish)")
    }
  }
  
  p1 <- DotPlot(object = x,
                group.by = cell_prediction,
                features = rev(marker_input),
                assay = "SCT",
                cols = c("#CC79A7","#009E73")) + 
    coord_flip() +
    ggtitle(title) +
    theme(axis.title = element_text(size = 20),
          axis.text.x = element_text(size=15, angle = 45, hjust=1),
          axis.text.y = element_text(size=15))
  
  #save
  
  pdf(out_file, height = 18, width = width_out)
  plot(p1)
  dev.off()
  
}, x = seurat_transcript_tmp, y = names(seurat_transcript_tmp))

rm(seurat_transcript_tmp)
```

# Gene/Transcript specific FeaturePlots

#####TODO: Change transcript feature plots to small dotplots (for the sake of space) in final figs (can keep this code just add the dotplot)


```{r}
genes_subset_pbmc <- c("CD8A", "CD79A")
genes_subset_lung <- c("C1QB", "CD3E")

# fetch all transcript for genes of interest

transcript_list_subset <- mapply(FUN = function(x,y) {
  
  if (y == "SC3pv3_transcript_isoquant" || y == "SC3pv3_transcript_oarfish") {
    features <- x[grep(paste(genes_subset_pbmc, collapse = "|"), x)]
    
    # order based on subset input
    features <- features[order(match(sub("\\..*$", "", features), genes_subset_pbmc))]
    
  } else {
    features <- x[grep(paste(genes_subset_lung, collapse = "|"), x)]
    
    # order based on subset input
    features <- features[order(match(sub("\\..*$", "", features), genes_subset_lung))]
  }
  
}, x = transcript_list, y = names(transcript_list))

# FeaturePlot (gene)
mapply(FUN = function(x, y) {
  
  if (y == "SC3pv3_gene_int") {
    cell_prediction <- "predicted.celltype.l2"
    out_file <- paste0(output_prefix, "/Featureplot_gene_level_PBMC.pdf")
    marker_input <- genes_subset_pbmc
    
  } else {
    cell_prediction <- "predicted.ann_level_3"
    out_file <- paste0(output_prefix, "/Featureplot_gene_level_lung.pdf")
    marker_input <- genes_subset_lung
  }
  
  # subset by scnanoseq only
  tmp_obj <- x
  Idents(tmp_obj) <- "sample_info"
  tmp_obj <- subset(tmp_obj, idents = "scnanoseq_Oxford")
  
  p1 <- FeaturePlot(object = tmp_obj,
                    features = marker_input,
                    cols = c("#D3D3D3", "#0072B2"),
                    ncol = 1)
  
  #save
  
  pdf(out_file, height = 12, width = 5)
  plot(p1)
  dev.off()
  
}, x = seurat_int, y = names(seurat_int))

# FeaturePlot (transcript)
mapply(FUN = function(x, y) {
  
  if (y == "SC3pv3_transcript_isoquant" || y == "SC3pv3_transcript_oarfish") {
    cell_prediction <- "predicted.celltype.l2"
    n = 3
    width_out <- 15
    marker_input <- transcript_list_subset[[y]]
    
    if (y == "SC3pv3_transcript_isoquant") {
      out_file <- paste0(output_prefix, "/Featureplot_transcript_level_PBMC_isoquant.pdf")
    } else {
      out_file <- paste0(output_prefix, "/Featureplot_transcript_level_PBMC_oarfish.pdf")
    }
    
  } else {
    cell_prediction <- "predicted.ann_level_3"
    marker_input <- transcript_list_subset[[y]]
    n = 4
    width_out <- 20
    
    if (y == "SC5pv2_transcript_isoquant") {
      out_file <- paste0(output_prefix, "/Featureplot_transcript_level_lung_isoquant.pdf")
    } else {
      out_file <- paste0(output_prefix, "/Featureplot_transcript_level_lung_oarfish.pdf")
    }
    
  }
  
  p1 <- FeaturePlot(object = x,
                    features = marker_input,
                    cols = c("#D3D3D3", "#009E73"),
                    ncol = n)
  
  #save
  
  pdf(out_file, height = 12, width = width_out)
  plot(p1)
  dev.off()
  
}, x = seurat_objs_transcript, y = names(seurat_objs_transcript))
```