---
title: "scnanoseq and BLAZE published (Ground Truth/You et al): figures"
author:
  - name: "Austyn Trull"
  - name: "Lara Ianov, Ph.D."
output:
  html_document:
    code_folding: hide
    css: "style.css"
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
set.seed(1234)

library(ggplot2)
library(Seurat)
library(dplyr)
library(grid)

# add custom functions from UBDS R scripts
lapply(list.files("./scripts/R"), FUN = function(x) source(paste0("./scripts/R/", x)))

#### CONSTANT DEFS ####

## The prefix to output files
output_prefix <- "./output_BLAZE_dataset_figures"

dir.create(file.path(output_prefix), recursive = TRUE)
```

# Load pre-processed Seurat objects

From our previous analysis for each dataset groups (original BLAZE matrices / You et al data and BLAZE data re-processed with `nf-core/scnanoseq`),
load both objects into the environment

```{r}
scnanoseq_objs <- readRDS(file = "./output_BLAZE_scnanoseq_data/seurat_scnanoseq_BLAZE.rds")
blaze_objs <- readRDS(file = "./output_BLAZE_published_data/seurat_BLAZE_ground_truth.rds")
```

# Subset data to PromethION dataset

Focus on main BLAZE figures are for the PromethION datasets. Subset the dataset for gene and transcript level data across both analysis

```{r}
scnanoseq_promith <- scnanoseq_objs[c(5,6,9)]
blaze_promith <- blaze_objs[c(2,5)]

scnanoseq_promith
blaze_promith
```

# Correlation between datasets for nFeature and nCount

For the cell barcodes from scnanoseq which are present in the blaze set, we compute the correlation between the datasets for nFeature and nCount (for gene and transcript-level ; thus 4 plots total)

```{r}
# merge metadata based on barcodes from scnanoseq

merged_meta <- mapply(FUN = function(x,y) {
  
  if (y == "ERR9958135_gene_isoquant") {
    blaze_meta <- blaze_promith$promith_gene@meta.data
  } else {
    blaze_meta <- blaze_promith$promith_trans@meta.data
  }
  
  if (y == "ERR9958135_transcript_oarfish") {
    suffix <- "_oarfish_scnanoseq"
  } else  {
    suffix <- "_isoquant_scnanoseq"
  }

  scnano_meta <- x@meta.data
  
  # add BCs as column
  scnano_meta$cell_barcodes <- rownames(scnano_meta)
  blaze_meta$cell_barcodes <- rownames(blaze_meta)
  
  merged_meta <- merge(x = scnano_meta,
                       y = blaze_meta,
                       by = "cell_barcodes",
                       suffixes= c(suffix, "_You_et_al"))
  
}, x=scnanoseq_promith, y=names(scnanoseq_promith), SIMPLIFY = FALSE)
```

```{r}
## nFeature scatters ##

# params

feature_type <- gsub(".*_(gene|transcript)_.*", "\\1", names(merged_meta))
output_feature <- paste0(gsub("^[^_]+_", "", names(merged_meta)),  "_nFeature.pdf")
feature_2 <- "nFeature_RNA_You_et_al"

mapply(FUN = function(x,y,z,out_file) {
  
  if (z == "ERR9958135_transcript_oarfish") {
    feature_1 <- "nFeature_RNA_oarfish_scnanoseq"
    axis_x <- paste0("Number of ", y, "s in oarfish scnanoseq")
    axis_y <- paste0("Number of ", y, "s in You et al.")
  } else if (z == "ERR9958135_transcript_isoquant") {
    feature_1 <- "nFeature_RNA_isoquant_scnanoseq"
    axis_x <- paste0("Number of ", y, "s in IsoQuant scnanoseq")
    axis_y <- paste0("Number of ", y, "s in You et al.")
  } else if (z == "ERR9958135_gene_isoquant") {
    feature_1 <- "nFeature_RNA_isoquant_scnanoseq"
    axis_x <- paste0("Number of ", y, "s in IsoQuant scnanoseq")
    axis_y <- paste0("Number of ", y, "s in You et al.")
  }
  
  scatter_plot <- scatter_corr(data.frame = x,
                               feature = feature_1,
                               feature2 = feature_2,
                               feature_type = y,
                               title_size = 30,
                               axis_size = 15,
                               point_size = 2,
                               point_alpha = 0.5) + 
    xlab(axis_x) +
    ylab(axis_y) +
    theme(axis.title = element_text(size = 20))
  
  file_name <- paste0(output_prefix, "/", out_file)

  #adding some room to the right to avoid axis being cutoff
  pdf(file = file_name, height = 8, width = 8)
  plot(scatter_plot + theme(plot.margin = margin(t = 20, b = 20, l = 50, r = 50)))
  dev.off()
  
}, x=merged_meta, y=feature_type, z = names(merged_meta), out_file = output_feature)

## nCount scatters ##

# params

feature_type <- gsub(".*_(gene|transcript)_.*", "\\1", names(merged_meta))
output_feature <- paste0(gsub("^[^_]+_", "", names(merged_meta)),  "_nCount.pdf")
feature_2 <- "nCount_RNA_You_et_al"

mapply(FUN = function(x,y,z, out_file) {
  
  if (z == "ERR9958135_transcript_oarfish") {
    feature_1 <- "nCount_RNA_oarfish_scnanoseq"
    axis_x <- paste0("Number of molecules in oarfish scnanoseq")
    axis_y <- paste0("Number of molecules in You et al.")
  } else {
    feature_1 <- "nCount_RNA_isoquant_scnanoseq"
    axis_x <- paste0("Number of molecules in IsoQuant scnanoseq")
    axis_y <- paste0("Number of molecules in You et al.") 
  }
  
  scatter_plot <- scatter_corr(data.frame = x,
                               feature = feature_1,
                               feature2 = feature_2,
                               feature_type = y,
                               title_size = 30,
                               axis_size = 15,
                               point_size = 2,
                               point_alpha = 0.5) + 
    xlab(axis_x) +
    ylab(axis_y) +
    theme(axis.title = element_text(size = 20))
  
  file_name <- paste0(output_prefix, "/", out_file)

  pdf(file = file_name, height = 8, width = 8)
  plot(scatter_plot + theme(plot.margin = margin(t = 20, b = 20, l = 50, r = 50)))
  dev.off()
  
}, x=merged_meta, y=feature_type, z = names(merged_meta), out_file = output_feature)
```

# UMAPS across PromethION datasets

```{r}
UMAPs <- mapply(FUN = function(x, y, z) {
  
  title_col <- ifelse(z == "gene", "#0072B2", "#009E73")
  
  p1 <- DimPlot(x,
                reduction = "umap",
                label = FALSE,
                pt.size = NULL) + 
    guides(color = guide_legend(override.aes = list(size = 4), ncol=1)) +
    ggtitle(paste0("scnanoseq: ", y)) +
    theme(legend.text = element_text(size = 15),
          axis.title = element_text(size = 20),
          axis.text.x = element_text(size=15),
          axis.text.y = element_text(size=15),
          plot.title = element_text(size = 25, color = title_col),
          strip.text.x = element_text(size = 20, face = "bold"))
  
  return(p1)
  
}, x = scnanoseq_promith, y = names(scnanoseq_promith), z = feature_type)


pdf(paste0(output_prefix, "/","UMAPS_scnanoseq.pdf"),
    width = 9,
    height = 12)

gridExtra::grid.arrange(grobs = UMAPs)

dev.off()


UMAPs <- mapply(FUN = function(x, y, z) {
  
  title_col <- ifelse(z == "gene", "#0072B2", "#009E73")
  
  p1 <- DimPlot(x,
                reduction = "umap",
                label = FALSE,
                pt.size = NULL) + 
    guides(color = guide_legend(override.aes = list(size = 6), ncol=1)) +
    ggtitle(paste0("You et al: ", y)) +
    theme(legend.text = element_text(size = 15),
          axis.title = element_text(size = 20),
          axis.text.x = element_text(size=15),
          axis.text.y = element_text(size=15),
          plot.title = element_text(size = 25, color = title_col),
          strip.text.x = element_text(size = 20, face = "bold"))
    
  
  return(p1)
  
}, x = blaze_promith, y = c("ERR9958135_gene", "ERR9958135_transcript"), z = c("gene", "transcript")) # using names from scnanoseq object for consistency in figures


pdf(paste0(output_prefix, "/","UMAPS_You_et_al.pdf"),
    width = 8,
    height = 8)

gridExtra::grid.arrange(grobs = UMAPs)

dev.off()
```

# nFeature UMAPs across PromethION datasets

```{r}
UMAPs <- mapply(FUN = function(x, y, z) {
  
  title_col <- ifelse(z == "gene", "#0072B2", "#009E73")

  p1 <- FeaturePlot(x,
                    reduction = "umap",
                    features = "nFeature_RNA",
                    label = FALSE,
                    pt.size = NULL) + 
    ggtitle(paste0("scnanoseq: ", y)) +
    theme(legend.text = element_text(size = 15),
          legend.title = element_text(size = 15),
          axis.title = element_text(size = 20),
          axis.text.x = element_text(size=15),
          axis.text.y = element_text(size=15),
          plot.title = element_text(size = 25, color = title_col)) +
    labs(color = "Feature count")
    
  
  return(p1)
  
}, x = scnanoseq_promith, y = names(scnanoseq_promith), z = feature_type)


pdf(paste0(output_prefix, "/","FeaturePlot_nFeature_scnanoseq.pdf"),
    width = 9,
    height = 12)

gridExtra::grid.arrange(grobs = UMAPs)

dev.off()


UMAPs <- mapply(FUN = function(x, y, z) {
  
  title_col <- ifelse(z == "gene", "#0072B2", "#009E73")

  p1 <- FeaturePlot(x,
                    reduction = "umap",
                    features = "nFeature_RNA",
                    label = FALSE,
                    pt.size = NULL) + 
    ggtitle(paste0("You et al: ", y)) +
    theme(legend.text = element_text(size = 15),
          legend.title = element_text(size = 15),
          axis.title = element_text(size = 20),
          axis.text.x = element_text(size=15),
          axis.text.y = element_text(size=15),
          plot.title = element_text(size = 25, color = title_col)) +
    labs(color = "Feature count")
  
  
  return(p1)
  
}, x = blaze_promith, y = c("ERR9958135_gene", "ERR9958135_transcript"), z = c("gene", "transcript"))


pdf(paste0(output_prefix, "/","FeaturePlot_nFeature_You_et_al.pdf"),
    width = 9,
    height = 12)

gridExtra::grid.arrange(grobs = UMAPs)

dev.off()
```
