---
title: "scnanoseq - 10X 3' and 5' dataset analysis"
author:
  - name: "Austyn Trull"
  - name: "Lara Ianov, Ph.D."
output:
  html_document:
    code_folding: hide
    css: "style.css"
    toc: true
    toc_float: true
---

This script performs standard post-secondary analysis on the 10X data 5' and 3' datasets discussed at the following application note: <https://www.10xgenomics.com/library/dea066>. The input data is the scnanoseq outputs (gene and transcript level) of both datasets, along with the Illumina data provided by 10X (ground-truth).

* Reference information: hg38 GENCODE v32 (Ensembl 98). This version matches the genome reference used in the Illumina analysis by 10X.

The input are the following files from the pipeline:

```
input/scnanoseq_10X_data/
├── SC3pv3_GEX_Human_PBMC_ONT.gene_counts.tsv
├── SC3pv3_GEX_Human_PBMC_ONT.transcript_counts.tsv
├── SC5pv2_GEX_Human_Lung_Carcinoma_DTC_ONT.gene_counts.tsv
└── SC5pv2_GEX_Human_Lung_Carcinoma_DTC_ONT.transcript_counts.tsv

input/CellRanger_10X_data/
├── SC3pv3_GEX_Human_PBMC_filtered_feature_bc_matrix.h5
└── SC5pv2_GEX_Human_Lung_Carcinoma_DTC_filtered_feature_bc_matrix.h5
```


```{r}
set.seed(1234)

library(ggplot2)
library(Seurat)
library(harmony)
library(dplyr)
library(tidyr)
library(grid)
library(clustree)
library(rtracklayer)
library(AnnotationHub)
library(Azimuth)
library(SeuratData)

# add custom functions from UBDS R scripts
lapply(list.files("./scripts/R"), FUN = function(x) source(paste0("./scripts/R/", x)))
```

```{r}
#### CONSTANT DEFS ####

## The prefix to output files
output_prefix <- "./output_scnanoseq_10X_data"

### --- scnanoseq Oxford data --- ###
# input files (sample and feature type [gene/transcript])
in_files_scnanoseq <- dir("./input/scnanoseq_10X_data",full.names = TRUE)
in_files_scnanoseq

# sample names
sample_names_scnanoseq <- dir("./input/scnanoseq_10X_data",full.names = FALSE)
sample_names_scnanoseq <- gsub("_.*","",sample_names_scnanoseq)

# feature
feature_type_scnanoseq <- dir("./input/scnanoseq_10X_data",full.names = FALSE)
feature_type_scnanoseq <- gsub(".*_ONT.","",feature_type_scnanoseq)

feature_type_scnanoseq <- gsub("_.*","",feature_type_scnanoseq)

feature_type_scnanoseq

# clean basenames with feature type
basenames_scnanoseq <- paste0(sample_names_scnanoseq, "_", feature_type_scnanoseq)
basenames_scnanoseq

### --- CellRanger (CR) Illumina data --- ###

in_files_CR <- dir("./input/CellRanger_10X_data",full.names = TRUE)
in_files_CR

# sample names
sample_names_CR <- dir("./input/CellRanger_10X_data",full.names = FALSE)
sample_names_CR <- gsub("_.*","",sample_names_CR)

# only gene-level here
feature_type_CR <- c("gene", "gene")

# clean basenames with feature type
basenames_CR <- paste0(sample_names_CR, "_", feature_type_CR, "_illumina")
basenames_CR
```

# PREPARE INPUT DATA

Import matrices, perform Ensembl to gene/transcript name conversion, make Seurat objects

```{r}
dir.create(file.path(output_prefix), recursive = TRUE)

# Read in the matrices
# At this time, gene and transcript will be part of the same list.
# If needed, feature type can be split at a later time
cell_bc_matrix <- lapply(setNames(in_files_scnanoseq,
                                 make.names(basenames_scnanoseq)),
               read.table, sep = "\t", header = TRUE, row.names = 1)
```

Convert Ensembl ids to gene/transcript names for scnanoseq data


```{r, eval=FALSE}
ah <- AnnotationHub()
```

```{r, eval=FALSE}
hg38_q <- query(ah, c("Ensembl", "Homo sapiens","98"))
hg38_q

hg38_q$title # finding the same gtf we used for generating 10X reference

hg38_annot <- subset(hg38_q, title == "Homo_sapiens.GRCh38.98.gtf")[[1]]
hg38_annot #info matches the gtf we used in CR_arc

# seqlevels are correct as we mapped to Ensembl genome
genome(hg38_annot) <- "hg38"
hg38_annot
```

```{r}
cell_bc_matrix_genes <- mapply(FUN = function(x,y,n) {
  
  if (y == "gene") {
    
    id_map <- data.frame(gene_id_version = rownames(x), # keep track of original
                         gene_id = gsub("\\.[0-9]+","", rownames(x)))
    
    # subset reference to metadata needed only
    
    hg38_subset <- as.data.frame(hg38_annot) %>%
      dplyr::filter(type == "gene") %>%
      dplyr::select(gene_id, gene_name)
    
    
    id_map <- id_map %>%
      plyr::join(as.data.frame(hg38_subset), by = c("gene_id")) %>%
      dplyr::mutate(gene_name = ifelse(is.na(gene_name), gene_id, gene_name)) #where gene_name == NA, use ID
    
    stopifnot(identical(id_map$gene_id_version, rownames(x)))
    
    rownames(x) <- make.names(id_map$gene_name, unique = TRUE)
    
    #save id_map
    
    file_name <- paste0(output_prefix, "/", names(cell_bc_matrix[n]), "_id_map.csv")
    
    write.csv(id_map, file = file_name, row.names = FALSE)
    
  } else {
    
    id_map <- data.frame(transcript_id_version = rownames(x), # keep track of original
                         transcript_id = gsub("\\.[0-9]+","",rownames(x)))
    
    # subset reference to metadata needed only
    
    hg38_subset <- as.data.frame(hg38_annot) %>%
      dplyr::filter(type == "transcript") %>%
      dplyr::select(transcript_id, transcript_name)
    
    
    id_map <- id_map %>%
      plyr::join(as.data.frame(hg38_subset), by = c("transcript_id")) %>%
      dplyr::mutate(transcript_name = ifelse(is.na(transcript_name), transcript_id, transcript_name))
    
    
    stopifnot(identical(id_map$transcript_id_version, rownames(x)))
    
    rownames(x) <- make.names(id_map$transcript_name, unique = TRUE)
    
    #save id_map
    
    file_name <- paste0(output_prefix, "/", names(cell_bc_matrix[n]), "_id_map.csv")
    
    write.csv(id_map, file = file_name, row.names = FALSE)
    
  }
  
  return(x)
  
}, x=cell_bc_matrix, y=feature_type_scnanoseq, n=1:length((cell_bc_matrix)),SIMPLIFY = FALSE)
```

```{r}
rm(cell_bc_matrix)
gc()
```

## Create Seurat objects

scnanoseq objects

```{r}
# Create the seurat object using RNA Assay
seurat_objs_scnanoseq <- mapply(FUN = function(x,y) {
  
  seurat_obj <- CreateSeuratObject(
    counts = x,
    assay = "RNA",
    min.cells = 1,
    min.features = 1,
    project = y)
  
  return(seurat_obj)
  
}, x=cell_bc_matrix_genes, y=names(cell_bc_matrix_genes))
```

CR objects

```{r}
CR_counts <- lapply(setNames(in_files_CR,
                             make.names(basenames_CR)),
                    Read10X_h5)

# Create the seurat object using RNA Assay
seurat_objs_CR <- mapply(FUN = function(x,y) {
  
  seurat_obj <- CreateSeuratObject(
    counts = x,
    assay = "RNA",
    min.cells = 1,
    min.features = 1,
    project = y)
  
  return(seurat_obj)
  
}, x=CR_counts, y=names(CR_counts))
```

Combining all objects into a single list to easily iterate over all until object integration

```{r}
seurat_objs <- c(seurat_objs_scnanoseq, seurat_objs_CR)

feature_type_all <- c(feature_type_scnanoseq, feature_type_CR)
```


# GENERATE QC'S

Mitochondrial % are computed based on feature biotype detected from Illumina data (given removal of other biotypes from the reference).

```{r, eval=FALSE}
# make vector of all gene names mapped to the mitochondrial chr.
hg38_mito_genes <- as.data.frame(hg38_annot) %>%
  filter(seqnames == "MT", type == "gene") %>%
  select(seqnames, gene_id, gene_name, gene_biotype)

hg38_mito_genes

hg38_mito_transcripts <- as.data.frame(hg38_annot) %>%
  filter(seqnames == "MT", type == "transcript") %>%
  select(seqnames, transcript_id, transcript_name, gene_biotype)

hg38_mito_transcripts
```


```{r, eval=FALSE}
# parse which feature names are present on a per sample basis
# feature names have been checked where general pattern usage below captures all biotypes
# note that scnanoseq patterns have `.` instead of `-` due to Ensembl to gene name conversion

hg38_mito_features <- mapply(FUN = function(x,y) {
  
  if (y == "gene") {
    
    hg38_mito_genes$in_data <- hg38_mito_genes$gene_name %in% rownames(x@assays$RNA$counts)
    
    # try for MT pattern with `.`
    if (all(hg38_mito_genes$in_data == FALSE)) {
      
      feauture_name <- gsub("-", ".", hg38_mito_genes$gene_name)
      
      hg38_mito_genes$gene_name <- feauture_name
      
      hg38_mito_genes$in_data <- feauture_name %in% rownames(x@assays$RNA$counts)
      
    }
    
    return(hg38_mito_genes)
    
  } else {
    
    hg38_mito_transcripts$in_data <- hg38_mito_transcripts$transcript_name %in% rownames(x@assays$RNA$counts)
    
    # try for MT pattern with `.`
    if (all(hg38_mito_transcripts$in_data == FALSE)) {
      
      feauture_name <- gsub("-", ".", hg38_mito_transcripts$transcript_name)
      
      hg38_mito_transcripts$transcript_name <- feauture_name
      
      hg38_mito_transcripts$in_data <- feauture_name %in% rownames(x@assays$RNA$counts)
      
    }
    
    return(hg38_mito_transcripts)
  }
  
}, x=seurat_objs, y=feature_type_all, SIMPLIFY = FALSE)

hg38_mito_features

# For any FALSE (genes not present in the dataset), filter out the genes
# Also filter by biotype to keep only protein_coding to match Illumina quantification
hg38_mito_features <- mapply(FUN = function(x,p) {
  
  sample_data_var <- paste0("in_data_", p)
  
  hg38_mito_features <- subset(x, in_data == TRUE & gene_biotype == "protein_coding")
  
  return(hg38_mito_features)
  
}, x=hg38_mito_features, p=names(hg38_mito_features), SIMPLIFY = FALSE)
```

Filter based on feature names detected above:

```{r}
# MITOCHONDRIAL PERCENT

seurat_objs <- mapply(FUN = function(x,y) {

  seurat_obj <- PercentageFeatureSet(object = x, features = y[,3], col.name = "percent.mt", assay = "RNA")
  
  return(seurat_obj)
  
}, x=seurat_objs, y=hg38_mito_features)

# brief look at the metadata

lapply(X = seurat_objs, FUN = function(x) {
    
    head(x@meta.data)

})
```

```{r}
# VIOLIN AND FEATURESCATTER, PRE-FILTER QC

mapply(FUN = function(x,y) {
  
  file_name1 <- paste0(output_prefix, "/", names(seurat_objs)[[y]], "_violin_plot_pre_filter.pdf")
  file_name2 <- paste0(output_prefix, "/", names(seurat_objs)[[y]], "_scatter_pre_filter.pdf")
  
  pdf(file = file_name1,
      width = 12,
      height = 6)
  
  plot(VlnPlot(object = x, features = c("nFeature_RNA", "nCount_RNA","percent.mt"), ncol = 3, cols = "#56B4E9", assay = "RNA"))
  
  dev.off()
  
  
  pdf(file = file_name2,
      width = 12,
      height = 6)
  
  plot1 <- FeatureScatter(object = x, feature1 = "nCount_RNA", feature2 = "percent.mt")
  plot2 <- FeatureScatter(object = x, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", cols = "#56B4E9")
  
  plot(plot1 + plot2) 
  
  dev.off()
  
  
}, x = seurat_objs, y = 1:length(seurat_objs))
```

```{r}
# Density plots per sample of nFeatures / nCount (unscaled)

all_plot_nFeatureRNA <- mapply(FUN = function(x) {
  
  p1 <- plotSingleCellDensity(input_obj = x,
                              metadata_variable = "nFeature_RNA",
                              group.by = "orig.ident",
                              scale_x_axis = FALSE) +
  	geom_vline(xintercept = 200)
  
  return(p1)

}, x = seurat_objs, SIMPLIFY = FALSE)

all_plot_nCountRNA <- mapply(FUN = function(x) {
  
  p1 <- plotSingleCellDensity(input_obj = x,
                              metadata_variable = "nCount_RNA",
                              group.by = "orig.ident",
                              scale_x_axis = FALSE)  +
  	geom_vline(xintercept = 300)
  
  return(p1)

}, x = seurat_objs, SIMPLIFY = FALSE)

all_plots_overview_nFeature <- cowplot::plot_grid(plotlist = all_plot_nFeatureRNA)
all_plots_overview_nCount <- cowplot::plot_grid(plotlist = all_plot_nCountRNA)

png(paste0(output_prefix, "/nFeature_un_scaled_RNA_pre_filter.png"), width = 1600, height = 1200)
plot(all_plots_overview_nFeature)
dev.off()

png(paste0(output_prefix, "/nCount_un_scaled_RNA_pre_filter.png"), width = 1600, height = 1200)
plot(all_plots_overview_nCount)
dev.off()
```

Scaled version

```{r, eval=FALSE}
# Density plots per sample of nFeatures / nCount (scaled)

all_plot_nFeatureRNA <- mapply(FUN = function(x) {
  
  p1 <- plotSingleCellDensity(input_obj = x,
                              metadata_variable = "nFeature_RNA",
                              group.by = "orig.ident",
                              scale_x_axis = TRUE) +
  	geom_vline(xintercept = 200)
  
  return(p1)

}, x = seurat_objs, SIMPLIFY = FALSE)

all_plot_nCountRNA <- mapply(FUN = function(x) {
  
  p1 <- plotSingleCellDensity(input_obj = x,
                              metadata_variable = "nCount_RNA",
                              group.by = "orig.ident",
                              scale_x_axis = TRUE) +
  	geom_vline(xintercept = 300)
  
  return(p1)

}, x = seurat_objs, SIMPLIFY = FALSE)

all_plots_overview_nFeature <- cowplot::plot_grid(plotlist = all_plot_nFeatureRNA)
all_plots_overview_nCount <- cowplot::plot_grid(plotlist = all_plot_nCountRNA)

png(paste0(output_prefix, "/nFeature_scaled_RNA_pre_filter.png"), width = 1600, height = 1200)
plot(all_plots_overview_nFeature)
dev.off()

png(paste0(output_prefix, "/nCount_scaled_RNA_pre_filter.png"), width = 1600, height = 1200)
plot(all_plots_overview_nCount)
dev.off()
```


```{r}
# Scatter Plot All 3 Variables

MT_preFil <- purrr::map_dfr(seurat_objs, ~ FetchData(.x, vars = c("orig.ident", "nCount_RNA", "nFeature_RNA",
                                       "percent.mt"))) %>% 
    ggplot(aes(x = nCount_RNA, y = nFeature_RNA, color = percent.mt)) +
  	geom_point() + 
  scale_colour_gradient(low = "grey90", high = "black", limits = c(0,100)) +
  	stat_smooth(method = lm) +
  	scale_x_log10() + 
  	scale_y_log10() + 
  	theme_classic() +
  	facet_wrap(~orig.ident, scales = "free") +
    theme_classic()

mt_prefilter_name <- paste0(output_prefix, "/mt_prefilter.png")
png(mt_prefilter_name, width = 1200, height = 800)
plot(MT_preFil)
dev.off()

# Save cell count prior to filtering
cell_count <- purrr::imap_dfr(seurat_objs, ~ FetchData(.x, vars = "orig.ident") %>% 
                                tally(name = "Cells_before_filtering") %>% 
                                mutate(Sample = Project(seurat_objs[[.y]]), .before = `Cells_before_filtering`))

write.csv(cell_count, paste0(output_prefix, "/cell_count_pre_filtering.csv"), row.names = FALSE)
```

Initial QCs suggest that there are several low quality cells, however, the overall number of cells displaying this metric are similar across methods.

# Correlation plots between gene-level scnanoseq and CR dataset (pre-filtering)

We generate these prior to any filtering, since filtering based on cutoff are bound to vary on technological differences and other factors (e.g. see note below).
Depending on this, the levels of filtration will vary (and downstream correlations will be limited to cells retained in both datasets. But here we show raw as well)

```{r}
# merge metadata based on barcodes from scnanoseq and CR runs (gene-level only)

merged_meta <- mapply(FUN = function(x,y) {
  
  scnano_meta <- x@meta.data
  CR_meta <- y@meta.data
  
  # add BCs as column
  scnano_meta$cell_barcodes <- rownames(scnano_meta)
  CR_meta$cell_barcodes <- rownames(CR_meta)
  
  # 10X CR contains -1 (data imported by keeping them for now, but need to remove for this step)
  
  CR_meta$cell_barcodes <- gsub("-1","",CR_meta$cell_barcodes)
  
  # merge by keeping all barcodes (not just what is common b/t the two)
  merged_meta <- merge(x = scnano_meta,
                       y = CR_meta,
                       by = "cell_barcodes",
                       suffixes= c("_scnanoseq", "_ground_truth"),
                       all = TRUE)
  
  # Change NA to 0 to display any barcodes unique to each dataset
  merged_meta <- merged_meta %>% mutate_if(is.numeric, ~replace_na(., 0))
  
}, x=seurat_objs[c(1,3)], y=seurat_objs[c(5,6)], SIMPLIFY = FALSE)
```

```{r}
## nFeature scatters ##

# params

feature_1 <- "nFeature_RNA_scnanoseq"
feature_2 <- "nFeature_RNA_ground_truth"
output_feature <- "_nFeature_pre_filter.png"

mapply(FUN = function(x,y) {
  
  scatter_plot <- scatter_corr(data.frame = x,
                               feature = feature_1,
                               feature2 = feature_2,
                               feature_type = "gene") # just genes here
  
  file_name <- paste0(output_prefix, "/", y, output_feature)

  #adding some room to the right to avoid axis being cutoff
  png(filename = file_name, height = 800, width = 1000)
  plot(scatter_plot + theme(plot.margin = margin(t = 20, b = 20, l = 50, r = 50)))
  dev.off()
  
}, x=merged_meta, y=names(merged_meta))

## nCount scatters ##

# params

feature_1 <- "nCount_RNA_scnanoseq"
feature_2 <- "nCount_RNA_ground_truth"
output_feature <- "_nCount_pre_filter.png"

mapply(FUN = function(x,y) {
  
  scatter_plot <- scatter_corr(data.frame = x,
                               feature = feature_1,
                               feature2 = feature_2,
                               feature_type = "gene")
  
  file_name <- paste0(output_prefix, "/", y, output_feature)

  png(filename = file_name, height = 800, width = 1000)
  plot(scatter_plot + theme(plot.margin = margin(t = 20, b = 20, l = 50, r = 50)))
  dev.off()
  
}, x=merged_meta, y=names(merged_meta))
```


# FILTERING

```{r}
#save object prior to filtering:
saveRDS(seurat_objs, paste0(output_prefix,"/all_10X_prefilter.rds"))
```


```{r}
# filtering criteria b/t gene and transcript is != based on QC plot (we do not need to set a higher cutoff for transcript-level)
nFeature_RNA_min <- c(400, 300, 400, 300, 400, 400)

# filter
seurat_objs <- mapply(FUN = function(x, y) {
  
  RNA_10X <- subset(x = x, subset= nFeature_RNA > y & nCount_RNA > 200 & percent.mt < 10)
  
}, x = seurat_objs, y = nFeature_RNA_min)
```

# POST-FILTER QC's

```{r}
# VIOLIN PLOTS

mapply(FUN = function(x,y) {
  
  file_name1 <- paste0(output_prefix, "/", names(seurat_objs)[[y]], "_violin_plot_post_filter.pdf")
  file_name2 <- paste0(output_prefix, "/", names(seurat_objs)[[y]], "_scatter_post_filter.pdf")
  
  pdf(file = file_name1,
      width = 12,
      height = 6)
  
  plot(VlnPlot(object = x, features = c("nFeature_RNA", "nCount_RNA","percent.mt"), ncol = 3, cols = "#56B4E9", assay = "RNA"))
  
  dev.off()
  
  
  pdf(file = file_name2,
      width = 12,
      height = 6)
  
  plot1 <- FeatureScatter(object = x, feature1 = "nCount_RNA", feature2 = "percent.mt")
  plot2 <- FeatureScatter(object = x, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", cols = "#56B4E9")
  
  plot(plot1 + plot2) 
  
  dev.off()
  
  
}, x = seurat_objs, y = 1:length(seurat_objs))
```

Scaled version

```{r, eval=FALSE}
# Density plots per sample of nFeatures / nCount (scaled)

all_plot_nFeatureRNA <- mapply(FUN = function(x) {
  
  p1 <- plotSingleCellDensity(input_obj = x,
                              metadata_variable = "nFeature_RNA",
                              group.by = "orig.ident",
                              scale_x_axis = TRUE)
  
  return(p1)

}, x = seurat_objs, SIMPLIFY = FALSE)

all_plot_nCountRNA <- mapply(FUN = function(x) {
  
  p1 <- plotSingleCellDensity(input_obj = x,
                              metadata_variable = "nCount_RNA",
                              group.by = "orig.ident",
                              scale_x_axis = TRUE)
  
  return(p1)

}, x = seurat_objs, SIMPLIFY = FALSE)

all_plots_overview_nFeature <- cowplot::plot_grid(plotlist = all_plot_nFeatureRNA)
all_plots_overview_nCount <- cowplot::plot_grid(plotlist = all_plot_nCountRNA)

png(paste0(output_prefix, "/nFeature_scaled_RNA_post_filter.png"), width = 1600, height = 1200)
plot(all_plots_overview_nFeature)
dev.off()

png(paste0(output_prefix, "/nCount_scaled_RNA_post_filter.png"), width = 1600, height = 1200)
plot(all_plots_overview_nCount)
dev.off()
```


```{r}
# Scatter Plot All 3 Variables

MT_preFil <- purrr::map_dfr(seurat_objs, ~ FetchData(.x, vars = c("orig.ident", "nCount_RNA", "nFeature_RNA",
                                       "percent.mt"))) %>% 
    ggplot(aes(x = nCount_RNA, y = nFeature_RNA, color = percent.mt)) +
  	geom_point() + 
  scale_colour_gradient(low = "grey90", high = "black", limits = c(0,100)) +
  	stat_smooth(method = lm) +
  	scale_x_log10() + 
  	scale_y_log10() + 
  	theme_classic() +
  	facet_wrap(~orig.ident, scales = "free") +
    theme_classic()

mt_prefilter_name <- paste0(output_prefix, "/mt_postfilter.png")
png(mt_prefilter_name, width = 1200, height = 800)
plot(MT_preFil)
dev.off()

# Save cell count prior to filtering
cell_count <- purrr::imap_dfr(seurat_objs, ~ FetchData(.x, vars = "orig.ident") %>% 
                                tally(name = "Cells_after_filtering") %>% 
                                mutate(Sample = Project(seurat_objs[[.y]]), .before = `Cells_after_filtering`))

write.csv(cell_count, paste0(output_prefix, "/cell_count_post_filtering.csv"), row.names = FALSE)
```

# Correlation plots between gene-level scnanoseq and CR dataset (post-filtering)

Unlike with the raw data, here we only investigate pattern across common cells retained due to differences in the effect that MT filtering is causing due to higher MT genes detected in Oxford vs Illumina.

```{r}
# merge metadata based on barcodes from scnanoseq and CR runs (gene-level only)

merged_meta <- mapply(FUN = function(x,y) {
  
  scnano_meta <- x@meta.data
  CR_meta <- y@meta.data
  
  # add BCs as column
  scnano_meta$cell_barcodes <- rownames(scnano_meta)
  CR_meta$cell_barcodes <- rownames(CR_meta)
  
  # 10X CR contains -1 (data imported by keeping them for now, but need to remove for this step)
  
  CR_meta$cell_barcodes <- gsub("-1","",CR_meta$cell_barcodes)
  
  # merge by keeping all barcodes (not just what is common b/t the two)
  merged_meta <- merge(x = scnano_meta,
                       y = CR_meta,
                       by = "cell_barcodes",
                       suffixes= c("_scnanoseq", "_ground_truth"))
  
  # Change NA to 0 to display any barcodes unique to each dataset
  merged_meta <- merged_meta %>% mutate_if(is.numeric, ~replace_na(., 0))
  
}, x=seurat_objs[c(1,3)], y=seurat_objs[c(5,6)], SIMPLIFY = FALSE)
```

```{r}
## nFeature scatters ##

# params

feature_1 <- "nFeature_RNA_scnanoseq"
feature_2 <- "nFeature_RNA_ground_truth"
output_feature <- "_nFeature_post_filter.png"

mapply(FUN = function(x,y) {
  
  scatter_plot <- scatter_corr(data.frame = x,
                               feature = feature_1,
                               feature2 = feature_2,
                               feature_type = "gene") # just genes here
  
  file_name <- paste0(output_prefix, "/", y, output_feature)

  #adding some room to the right to avoid axis being cutoff
  png(filename = file_name, height = 800, width = 1000)
  plot(scatter_plot + theme(plot.margin = margin(t = 20, b = 20, l = 50, r = 50)))
  dev.off()
  
}, x=merged_meta, y=names(merged_meta))

## nCount scatters ##

# params

feature_1 <- "nCount_RNA_scnanoseq"
feature_2 <- "nCount_RNA_ground_truth"
output_feature <- "_nCount_post_filter.png"

mapply(FUN = function(x,y) {
  
  scatter_plot <- scatter_corr(data.frame = x,
                               feature = feature_1,
                               feature2 = feature_2,
                               feature_type = "gene")
  
  file_name <- paste0(output_prefix, "/", y, output_feature)

  png(filename = file_name, height = 800, width = 1000)
  plot(scatter_plot + theme(plot.margin = margin(t = 20, b = 20, l = 50, r = 50)))
  dev.off()
  
}, x=merged_meta, y=names(merged_meta))
```

# `SCTransform` v2

`SCTransform` v2 is applied for the normalization of RNA data 

```{r}
seurat_objs <- mapply(FUN = function(x) {
  
  seurat_obj <- SCTransform(x,
                            assay = "RNA",
                            vars.to.regress = "percent.mt",
                            verbose = TRUE,
                            vst.flavor = "v2",
                            method = "glmGamPoi")
  
  return(seurat_obj)
  
}, x=seurat_objs)
```

Change to SCT assay:

```{r}
seurat_objs <- lapply(X = seurat_objs, FUN = function(x) {
    
    DefaultAssay(x) <- "SCT"
    
    return(x)
    
})
```

# Harmony integration (gene-level)

For the gene-level datasets, we integrate the Oxford and Illumina data for 3' and 5' datasets, creating 2 integrated objects across technologies/datasets.

```{r}
# subset per sample object, to only datasets to be integrated (gene-level only)

seurat_objs_3_prime <- seurat_objs[c("SC3pv3_gene", "SC3pv3_gene_illumina")]
seurat_objs_5_prime <- seurat_objs[c("SC5pv2_gene", "SC5pv2_gene_illumina")]
```

```{r}
# ref: https://github.com/satijalab/seurat/issues/4896

### 3 prime integration ###

int_features_3_prime <- SelectIntegrationFeatures(seurat_objs_3_prime, nfeatures = 3000)

# merge datasets prior to harmony and add integration features
seurat_objs_3_prime_int <- merge(x = seurat_objs_3_prime[[1]],
                                 y = seurat_objs_3_prime[2:length(seurat_objs_3_prime)],
                                 merge.data=TRUE)

VariableFeatures(seurat_objs_3_prime_int) <- int_features_3_prime

#PCA
seurat_objs_3_prime_int <- RunPCA(seurat_objs_3_prime_int, npcs = 50)

#orig.ident correspond to Illumina or Oxford sequencing
seurat_objs_3_prime_int <- RunHarmony(seurat_objs_3_prime_int,
                                      assay.use="SCT",
                                      group.by.vars = "orig.ident")

### 5 prime integration ###

int_features_5_prime <- SelectIntegrationFeatures(seurat_objs_5_prime, nfeatures = 3000)

# merge datasets prior to harmony and add integration features
seurat_objs_5_prime_int <- merge(x = seurat_objs_5_prime[[1]],
                                 y = seurat_objs_5_prime[2:length(seurat_objs_5_prime)],
                                 merge.data=TRUE)

VariableFeatures(seurat_objs_5_prime_int) <- int_features_5_prime

#PCA
seurat_objs_5_prime_int <- RunPCA(seurat_objs_5_prime_int, npcs = 50)

#orig.ident correspond to Illumina or Oxford sequencing
seurat_objs_5_prime_int <- RunHarmony(seurat_objs_5_prime_int,
                                      assay.use="SCT",
                                      group.by.vars = "orig.ident")
```

Save integrated objects:

```{r}
# creating a list of integrated objects:

seurat_int <- list("SC3pv3_gene_int" = seurat_objs_3_prime_int,
                   "SC5pv2_gene_int" = seurat_objs_5_prime_int)

saveRDS(seurat_int, paste0(output_prefix,"/seurat_integrated_objs.rds"))
```

# Azimuth

First slimming down the object as Azimuth performs normalization internally on RNA assay, then run Azimuth and save metadata to be imported into objects

```{r}
# set `.libPaths` for azimuth reference to go into getwd instead of container's home dir
.libPaths(getwd())

# will create a separate obj for azimuth process and add meteadat to main object later on

seurat_int_azimuth <- seurat_int

# see available datasets
available_data <- AvailableData()
available_data[grep("Azimuth", available_data[, 3]), 1:3] #pbmcref fo 3' and lungref for 5'

azimut_ref <- c("pbmcref", "lungref")

seurat_int_azimuth <- mapply(FUN = function(x,y,z) {
  
  seurat_azimuth <- RunAzimuth(query = x,
                               reference = y,
                               assay = "RNA")


}, x=seurat_int_azimuth, y=azimut_ref)

# Plot prelim azimuth annotations on each level 
# (reminder this is not on harmony UMAP as azimuth performs separate steps - we import metadata into ours):

p1 <- DimPlot(seurat_int_azimuth$SC3pv3_gene_int, group.by = "predicted.celltype.l1", label = TRUE, label.size = 3)
p2 <- DimPlot(seurat_int_azimuth$SC3pv3_gene_int, group.by = "predicted.celltype.l2", label = TRUE, label.size = 3)
p3 <- DimPlot(seurat_int_azimuth$SC3pv3_gene_int, group.by = "predicted.celltype.l3",label = TRUE, label.size = 3)

all_predicted <- cowplot::plot_grid(plotlist = list(p1,p2,p3))

png(file=paste0(output_prefix, "/predicted_cell_types_azimuth_pbmc.png"), width = 1600, height = 1200)
plot(all_predicted)
dev.off()

p1 <- DimPlot(seurat_int_azimuth$SC5pv2_gene_int, group.by = "predicted.ann_level_1", label = TRUE, label.size = 3)
p2 <- DimPlot(seurat_int_azimuth$SC5pv2_gene_int, group.by = "predicted.ann_level_2", label = TRUE, label.size = 3)
p3 <- DimPlot(seurat_int_azimuth$SC5pv2_gene_int, group.by = "predicted.ann_level_3",label = TRUE, label.size = 3)
p4 <- DimPlot(seurat_int_azimuth$SC5pv2_gene_int, group.by = "predicted.ann_level_4",label = TRUE, label.size = 3)
p5 <- DimPlot(seurat_int_azimuth$SC5pv2_gene_int, group.by = "predicted.ann_level_5",label = TRUE, label.size = 3)
p6 <- DimPlot(seurat_int_azimuth$SC5pv2_gene_int, group.by = "predicted.ann_finest_level",label = TRUE, label.size = 3)

all_predicted <- cowplot::plot_grid(plotlist = list(p1,p2,p3,p4,p5,p6))

png(file=paste0(output_prefix, "/predicted_cell_types_azimuth_lung.png"), width = 2000, height = 1200)
plot(all_predicted)
dev.off()
```

Add Azimuth metadata to main integrated object and save azimuth

```{r}
# add metadata to main object
seurat_int <- mapply(FUN = function(x, y) {
  
  # trim down metadata to variables desired to be added to main obj
  
  meta <- as.data.frame(y@meta.data)
  
  meta <- dplyr::select(meta, starts_with(c("predicted.", "mapping.")))
  
  seurat_obj <- AddMetaData(x, meta)
  
  return(seurat_obj)
  
}, x=seurat_int, y=seurat_int_azimuth)

lapply(X = seurat_int, FUN = function(x) {
    
    head(x@meta.data)

})

# save

saveRDS(seurat_int_azimuth, paste0(output_prefix,"/seurat_azimuth.rds"))
```

# DIMENSIONS AND RESOLUTIONS (gene-level)

## ELBOW PLOT

```{r}
mapply(FUN = function(x, y) {
  
  elbow_plot <- paste0(output_prefix, "/", y , "_elbow.png")
  
  png(elbow_plot,
      width = 480,
      height = 480)
  
  plot(ElbowPlot(object = x,
                 ndims = 50,
                 reduction = "pca"))
  
  dev.off()
  
}, x = seurat_int, y = names(seurat_int))
```

## DIMS and RES (gene-level) 

The loop below will cycle through a set of dimensions (based on elbow plot) and resolutions to aid in a sensible choice of dims and resolution for the integrated object. A temporary object is created for the following process

```{r}
dim_output_dir <- paste0(output_prefix,"/dims_res_explore/")

dir.create(dim_output_dir, recursive = TRUE)

seurat_obj_tmp <- seurat_int

cluster_res <- seq(from = 0.1, to = 0.5, by = 0.1)

#NOTE: for future validations, can focus on higher dims (move range up)
dim <- seq(from = 15, to = 25, by = 1)

res <- tidyr::expand_grid(cluster_res, dim)

for (obj in names(seurat_obj_tmp) ) {
  
  out_path <- paste0(dim_output_dir, obj)
  
  dir.create(out_path, recursive = TRUE, showWarnings = FALSE)
  
  for (i in seq(nrow(res))) {
    
    seurat_obj_tmp[[obj]] <- RunUMAP(seurat_obj_tmp[[obj]],
                                     reduction = "harmony", 
                                     dims = 1:res[i,]$dim)
    
    seurat_obj_tmp[[obj]] <- FindNeighbors(seurat_obj_tmp[[obj]],
                                           reduction = "harmony",
                                           dims = 1:res[i,]$dim)
    
    seurat_obj_tmp[[obj]] <- FindClusters(seurat_obj_tmp[[obj]],
                                          resolution = res[i,]$cluster_res)
    
    file_name <- paste0(out_path,"/",
                        res[i,]$dim,
                        "_res_",
                        res[i,]$cluster_res,
                        ".png")
    
    p1 <- DimPlot(seurat_obj_tmp[[obj]], label = TRUE) + 
      ggtitle(paste0("Dim_res: ", obj, "_",res[i,]$dim, "_", res[i,]$cluster_res))
    
    # plot second/third level of azimuth predicted cell types
    
    if (obj == "SC3pv3_gene_int") {
      cell_prediction <- "predicted.celltype.l2"
    } else {
      cell_prediction <- "predicted.ann_level_3"
    }
    
    p2 <- DimPlot(seurat_obj_tmp[[obj]], group.by = cell_prediction, label = TRUE)
    
    dim_res_plot <- p1 + p2
    
    png(file_name,
        width = 1200,
        height = 800)
    
    plot(dim_res_plot)
    
    dev.off()
    
  }
}

rm(seurat_obj_tmp)
gc()
```

## CLUSTREE (gene-level)

For specific chosen dimension, generate clustree plot(s)

```{r}
dim <- 25

tmp_seurat_clustree <- seurat_int

clustree_res <- seq(from = 0, to = 0.7, by = 0.1)

res <- tidyr::expand_grid(clustree_res, dim)

for (obj in names(tmp_seurat_clustree) ) {
  
  out_path <- paste0(dim_output_dir, obj)
  
  for (i in seq(nrow(res))) {
    tmp_seurat_clustree[[obj]] <- RunUMAP(tmp_seurat_clustree[[obj]], 
                                          reduction = "harmony",
                                          dims = 1:dim)
    
    tmp_seurat_clustree[[obj]] <- FindNeighbors(tmp_seurat_clustree[[obj]],
                                                reduction = "harmony",
                                                dims = 1:dim)
    
    tmp_seurat_clustree[[obj]] <- FindClusters(tmp_seurat_clustree[[obj]],
                                               resolution = res[i,]$clustree_res)
  }
  
  # for each object, generate clustree map
  
  clustree_plot <- clustree(tmp_seurat_clustree[[obj]])
  
  pdf(paste0(out_path, "/",
             obj,"_", 
             "clustree.dim_", dim, ".pdf"),
      width = 8,
      height = 12)
  plot(clustree_plot)
  dev.off()
  
}

rm(tmp_seurat_clustree)
gc()
```

## CLUSTER (gene-level)

With appropriate dim and res, cluster the objects:
Note: given scope of the analysis and use of azimuth based annotation, this choice can be an approximate
to the azimith clusters. Further refinement can be performed for other types of downstream analysis.

```{r}
final_dim <- 25
final_res <- c(0.6, 0.5)

seurat_int <- mapply(FUN = function(x,y) {
  
  seurat_obj <- RunUMAP(obj = x,
                        reduction = "harmony",
                        dims = 1:final_dim)
  
  seurat_obj <- FindNeighbors(obj = seurat_obj,
                              reduction = "harmony",
                              dims = 1:final_dim)
  
  seurat_obj <- FindClusters(obj = seurat_obj,
                             resolution = y)
  
  return(seurat_obj)
  
}, x = seurat_int, y = final_res)
```

## POST-INTEGRATION PLOTS (gene-level)

### UMAPS with predicted cell types
```{r}
all_UMAPS <- mapply(FUN = function(x, y) {
  
  p1 <- DimPlot(x,
                reduction = "umap",
                label = TRUE,
                pt.size = 1) + 
    guides(color = guide_legend(override.aes = list(size = 8), ncol=1)) +
    theme(legend.text = element_text(size = 30)) +
    ggtitle(paste0("Gene-level: ", y))
  
  if (y == "SC3pv3_gene_int") {
    cell_prediction <- "predicted.celltype.l2"
  } else {
    cell_prediction <- "predicted.ann_level_3"
  }
  
  p2 <- DimPlot(x,
                group.by = cell_prediction,
                reduction = "umap",
                label = TRUE,
                pt.size = 1) + 
    guides(color = guide_legend(override.aes = list(size = 8), ncol=1)) +
    theme(legend.text = element_text(size = 30))
  
  return(p1 + p2)
  
}, x = seurat_int, y = names(seurat_int))


png(paste0(output_prefix, "/","All_UMAPS_pbmc.png"),
    width = 2000,
    height = 1200)

plot(all_UMAPS$SC3pv3_gene_int)

dev.off()

png(paste0(output_prefix, "/","All_UMAPS_lung.png"),
    width = 2000,
    height = 1200)


plot(all_UMAPS$SC5pv2_gene_int)

dev.off()
```

### UMAPS split by technology/pipeline

```{r}
all_UMAPS <- mapply(FUN = function(x, y) {
  
  p1 <- DimPlot(x,
                reduction = "umap",
                split.by = "orig.ident",
                label = TRUE,
                pt.size = 1) + 
    guides(color = guide_legend(override.aes = list(size = 8), ncol=1)) +
    theme(legend.text = element_text(size = 30)) +
    ggtitle(paste0("Gene-level: ", y))
  
  if (y == "SC3pv3_gene_int") {
    cell_prediction <- "predicted.celltype.l2"
  } else {
    cell_prediction <- "predicted.ann_level_3"
  }
  
  p2 <- DimPlot(x,
                group.by = cell_prediction,
                split.by = "orig.ident",
                reduction = "umap",
                label = TRUE,
                pt.size = 1) + 
    guides(color = guide_legend(override.aes = list(size = 8), ncol=1)) +
    theme(legend.text = element_text(size = 30))
  
  return(p1 + p2)
  
}, x = seurat_int, y = names(seurat_int))


png(paste0(output_prefix, "/All_UMAPS_pbmc_split.png"),
    width = 2200,
    height = 1200)

plot(all_UMAPS$SC3pv3_gene_int)

dev.off()

png(paste0(output_prefix, "/All_UMAPS_lung_split.png"),
    width = 2200,
    height = 1200)


plot(all_UMAPS$SC5pv2_gene_int)

dev.off()
```

### Post-integration QCs (post cell labeling ; gene-level)

```{r}
dir.create(paste0(output_prefix, "/post_label_QCs"), recursive = TRUE)

mapply(FUN = function(x, y) {
  
  if (y == "SC3pv3_gene_int") {
    cell_prediction <- "predicted.celltype.l2"
  } else {
    cell_prediction <- "predicted.ann_level_3"
  }
  
  p1 <- plotCounts(input_obj = x,
                   metadata_variable = cell_prediction,
                   group.by = "orig.ident") +
    theme(axis.text = element_text(size = 30))
  
  p2 <- plotCountsProportions(input_obj = x,
                              metadata_variable = cell_prediction,
                              group.by = "orig.ident") + 
    theme(axis.text.x = element_text(angle = 90, size = 30),
          axis.text.y = element_text(size = 30))
  
  
  all_plots <- cowplot::plot_grid(plotlist = list(p1, p2), ncol = 1,
                                  labels = "AUTO")
  
  png(paste0(output_prefix, "/post_label_QCs/Cell_counts_proportions", y, ".png"),width=1400,height=1600)
  plot(all_plots)
  dev.off()
  
}, x = seurat_int, y = names(seurat_int))

saveRDS(seurat_int, paste0(output_prefix,"/seurat_int_gene_level.rds"))
```

# DIMENSIONS AND RESOLUTIONS (transcript-level)

## PCA AND ELBOW PLOT

```{r}
seurat_objs_transcript <- seurat_objs[c("SC3pv3_transcript", "SC5pv2_transcript")]

# Add azimuth metadata and compute how many cells contain this metadata in transcript datasets

# add metadata to main object
seurat_objs_transcript <- mapply(FUN = function(x, y) {
  
  # trim down metadata to variables desired to be added to main obj
  
  meta <- as.data.frame(y@meta.data)
  
  meta <- dplyr::select(meta, starts_with(c("predicted.", "mapping.")))
  
  seurat_obj <- AddMetaData(x, meta)
  
  return(seurat_obj)
  
}, x=seurat_objs_transcript, y=seurat_int_azimuth)

lapply(X = seurat_objs_transcript, FUN = function(x) {
    
    head(x@meta.data)

})

# choose one prediction and compute ratio of how many cells contain azimuth annotation:
mapply(FUN = function(x, y) {
  
  if (y == "SC3pv3_transcript") {
    cell_prediction <- "predicted.celltype.l2"
  } else {
    cell_prediction <- "predicted.ann_level_3"
  }
  
  print(table(is.na(x@meta.data[[cell_prediction]])))
  
  print("Ratio of barcodes containing Azimith annotations is:") # FALSE mean it has metadata
  
  cell_with_meta <- as.data.frame(table(is.na(x@meta.data[[cell_prediction]])))$Freq[1]
  
  print((cell_with_meta / nrow(x@meta.data)) * 100)

}, x = seurat_objs_transcript, y = names(seurat_objs_transcript))
```

```{r}
seurat_objs_transcript <- mapply(FUN = function(x) {
  
  seurat_obj <- RunPCA(x, npcs = 50)
  
  return(seurat_obj)
  
}, x=seurat_objs_transcript)


mapply(FUN = function(x, y) {
  
  elbow_plot <- paste0(output_prefix, "/", y , "_elbow.png")
  
  png(elbow_plot,
      width = 480,
      height = 480)
  
  plot(ElbowPlot(object = x,
                 ndims = 50,
                 reduction = "pca"))
  
  dev.off()
  
}, x = seurat_objs_transcript, y = names(seurat_objs_transcript))
```

## DIMS and RES (gene-level) 

The loop below will cycle through a set of dimensions (based on elbow plot) and resolutions to aid in a sensible choice of dims and resolution for the integrated object. A temporary object is created for the following process

```{r}
dim_output_dir <- paste0(output_prefix,"/dims_res_explore/")

dir.create(dim_output_dir, recursive = TRUE)

seurat_obj_tmp <- seurat_objs_transcript

cluster_res <- seq(from = 0.1, to = 0.6, by = 0.1)

dim <- seq(from = 20, to = 27, by = 1)

res <- tidyr::expand_grid(cluster_res, dim)

for (obj in names(seurat_obj_tmp) ) {
  
  out_path <- paste0(dim_output_dir, obj)
  
  dir.create(out_path, recursive = TRUE, showWarnings = FALSE)
  
  for (i in seq(nrow(res))) {
    
    seurat_obj_tmp[[obj]] <- RunUMAP(seurat_obj_tmp[[obj]],
                                     reduction = "pca", 
                                     dims = 1:res[i,]$dim)
    
    seurat_obj_tmp[[obj]] <- FindNeighbors(seurat_obj_tmp[[obj]],
                                           reduction = "pca",
                                           dims = 1:res[i,]$dim)
    
    seurat_obj_tmp[[obj]] <- FindClusters(seurat_obj_tmp[[obj]],
                                          resolution = res[i,]$cluster_res)
    
    file_name <- paste0(out_path,"/",
                        res[i,]$dim,
                        "_res_",
                        res[i,]$cluster_res,
                        ".png")
    
    p1 <- DimPlot(seurat_obj_tmp[[obj]], label = TRUE) + 
      ggtitle(paste0("Dim_res: ", obj, "_",res[i,]$dim, "_", res[i,]$cluster_res))
    
    # plot second/third level of azimuth predicted cell types
    
    if (obj == "SC3pv3_transcript") {
      cell_prediction <- "predicted.celltype.l2"
    } else {
      cell_prediction <- "predicted.ann_level_3"
    }
    
    p2 <- DimPlot(seurat_obj_tmp[[obj]], group.by = cell_prediction, label = TRUE)
    
    dim_res_plot <- p1 + p2
    
    png(file_name,
        width = 1200,
        height = 800)
    
    plot(dim_res_plot)
    
    dev.off()
    
  }
}

rm(seurat_obj_tmp)
gc()
```


## CLUSTREE (gene-level)

For specific chosen dimension, generate clustree plot(s)

```{r}
dim <- 25

tmp_seurat_clustree <- seurat_objs_transcript

clustree_res <- seq(from = 0, to = 0.8, by = 0.1)

res <- tidyr::expand_grid(clustree_res, dim)

for (obj in names(tmp_seurat_clustree) ) {
  
  out_path <- paste0(dim_output_dir, obj)
  
  for (i in seq(nrow(res))) {
    tmp_seurat_clustree[[obj]] <- RunUMAP(tmp_seurat_clustree[[obj]], 
                                          reduction = "pca",
                                          dims = 1:dim)
    
    tmp_seurat_clustree[[obj]] <- FindNeighbors(tmp_seurat_clustree[[obj]],
                                                reduction = "pca",
                                                dims = 1:dim)
    
    tmp_seurat_clustree[[obj]] <- FindClusters(tmp_seurat_clustree[[obj]],
                                               resolution = res[i,]$clustree_res)
  }
  
  # for each object, generate clustree map
  
  clustree_plot <- clustree(tmp_seurat_clustree[[obj]])
  
  pdf(paste0(out_path, "/",
             obj,"_", 
             "clustree.dim_", dim, ".pdf"),
      width = 8,
      height = 12)
  plot(clustree_plot)
  dev.off()
  
}

rm(tmp_seurat_clustree)
gc()
```

## CLUSTER (transcript-level)

With appropriate dim and res, cluster the objects:

```{r}
final_dim <- 25
final_res <- c(0.8, 0.6)

seurat_objs_transcript <- mapply(FUN = function(x,y) {
  
  seurat_obj <- RunUMAP(obj = x,
                        reduction = "pca",
                        dims = 1:final_dim)
  
  seurat_obj <- FindNeighbors(obj = seurat_obj,
                              reduction = "pca",
                              dims = 1:final_dim)
  
  seurat_obj <- FindClusters(obj = seurat_obj,
                             resolution = y)
  
  return(seurat_obj)
  
}, x = seurat_objs_transcript, y = final_res)
```

## POST-INTEGRATION PLOTS (transcript-level)

### UMAPS with predicted cell types
```{r}
all_UMAPS <- mapply(FUN = function(x, y) {
  
  p1 <- DimPlot(x,
                reduction = "umap",
                label = TRUE,
                pt.size = 1) + 
    guides(color = guide_legend(override.aes = list(size = 8), ncol=1)) +
    theme(legend.text = element_text(size = 30)) +
    ggtitle(paste0("Transcript-level: ", y))
  
  if (y == "SC3pv3_transcript") {
    cell_prediction <- "predicted.celltype.l2"
  } else {
    cell_prediction <- "predicted.ann_level_3"
  }
  
  p2 <- DimPlot(x,
                group.by = cell_prediction,
                reduction = "umap",
                label = TRUE,
                pt.size = 1) + 
    guides(color = guide_legend(override.aes = list(size = 8), ncol=1)) +
    theme(legend.text = element_text(size = 30))
  
  return(p1 + p2)
  
}, x = seurat_objs_transcript, y = names(seurat_objs_transcript))


png(paste0(output_prefix, "/","All_UMAPS_pbmc_transcript.png"),
    width = 2000,
    height = 1200)

plot(all_UMAPS$SC3pv3_transcript)

dev.off()

png(paste0(output_prefix, "/","All_UMAPS_lung_transcript.png"),
    width = 2000,
    height = 1200)


plot(all_UMAPS$SC5pv2_transcript)

dev.off()
```


```{r}
saveRDS(seurat_objs_transcript, paste0(output_prefix,"/seurat_objs_transcript_level.rds"))
```

### Additional figures QCs (post cell labeling ; transcript-level)

```{r}
dir.create(paste0(output_prefix, "/post_label_QCs"), recursive = TRUE)

# just counts, no proportion here
mapply(FUN = function(x, y) {
  
  if (y == "SC3pv3_transcript") {
    cell_prediction <- "predicted.celltype.l2"
  } else {
    cell_prediction <- "predicted.ann_level_3"
  }
  
  p1 <- plotCounts(input_obj = x,
                   metadata_variable = cell_prediction,
                   group.by = "orig.ident") +
    theme(axis.text = element_text(size = 30))

  png(paste0(output_prefix, "/post_label_QCs/Cell_counts_proportions", y, ".png"),width=1400,height=800)
  plot(p1)
  dev.off()
  
}, x = seurat_objs_transcript, y = names(seurat_objs_transcript))
```

#### Feature-level figures (gene and transcript)

Prelim exploration of gene-level vs transcript-level expression for azimith markers

```{r}
dir.create(paste0(output_prefix, "/feature_expression"), recursive = TRUE)

# import feature names associated with predicted cells types

# PBMC FEATURES ###
pbmc_features <- read.csv("./input/references/azimuth_pbmb_celltype_l2_features.csv")

# combine all features for now
pbmc_features <- unlist(strsplit(pbmc_features$Markers, ", "))

pbmc_features <- unique(pbmc_features)

# search for all transcript under a gene name

pbmc_trascripts <- list()

# Searches for transcript features within SCT counts
pbmc_trascripts <- mapply(FUN = function(x) {
  
  feature <- grep(pattern = paste0("^", x, "\\."),
                  x = rownames(seurat_objs_transcript$SC3pv3_transcript@assays$SCT@counts),
                  value = TRUE)
  
}, x = pbmc_features, SIMPLIFY = FALSE)

pbmc_trascripts <- pbmc_trascripts[lengths(pbmc_trascripts) > 0]

### LUNG FEATURES ###

lung_features <- read.csv("./input/references/azimuth_lung_ann_level_3_features.csv")

# combine all features for now
lung_features <- unlist(strsplit(lung_features$Markers, ", "))

lung_features <- unique(lung_features)

# search for all transcript under a gene name

lung_transcripts <- list()

# Searches for transcript features within SCT counts
lung_transcripts <- mapply(FUN = function(x) {
  
  feature <- grep(pattern = paste0("^", x, "\\."),
                  x = rownames(seurat_objs_transcript$SC5pv2_transcript@assays$SCT@counts),
                  value = TRUE)
  
}, x = lung_features, SIMPLIFY = FALSE)

lung_transcripts <- lung_transcripts[lengths(lung_transcripts) > 0]

### gene ###

mapply(FUN = function(x) {
  
  if (x %in% rownames(seurat_int$SC3pv3_gene_int@assays$SCT@counts)) {
    
    p1 <- FeaturePlot(object = seurat_int$SC3pv3_gene_int,
                      split.by = "orig.ident", # split by technology
                      cols = c("lightgrey", "#0072B2"),
                      features = x, reduction = "umap",
                      keep.scale = "all")
    
    png(paste0(output_prefix, "/feature_expression/",x, "_pbmc_gene.png"),width=700,height=500)
    plot(p1)
    dev.off()
    
  } else {
    
    message(paste0(x, " not present in object. Skipping it."))
    
  }
  
}, x = pbmc_features)


mapply(FUN = function(x) {
  
  if (x %in% rownames(seurat_int$SC5pv2_gene_int@assays$SCT@counts)) {
    
    p1 <- FeaturePlot(object = seurat_int$SC5pv2_gene_int,
                      split.by = "orig.ident", # split by technology
                      cols = c("lightgrey", "#0072B2"),
                      features = x, reduction = "umap",
                      keep.scale = "all")
    
    png(paste0(output_prefix, "/feature_expression/",x, "_lung_gene.png"),width=700,height=500)
    plot(p1)
    dev.off()
    
  } else {
    
    message(paste0(x, " not present in object. Skipping it."))
    
  }
  
}, x = lung_features)

### transcript ###

mapply(FUN = function(x,y) {
  
  p1 <- FeaturePlot(object = seurat_objs_transcript$SC3pv3_transcript,
                    cols = c("lightgrey", "#0072B2"),
                    features = x, reduction = "umap",
                    keep.scale = "all") # all for visualization across isoforms
  
  png(paste0(output_prefix, "/feature_expression/",y, "_pbmc_transcript.png"),width=1500,height=1200)
  plot(p1)
  dev.off()
  
}, x = pbmc_trascripts, y = names(pbmc_trascripts))

mapply(FUN = function(x,y) {
  
  p1 <- FeaturePlot(object = seurat_objs_transcript$SC5pv2_transcript,
                    cols = c("lightgrey", "#0072B2"),
                    features = x, reduction = "umap",
                    keep.scale = "all") # all for visualization across isoforms
  
  png(paste0(output_prefix, "/feature_expression/",y, "_lung_transcript.png"),width=1800,height=1200)
  plot(p1)
  dev.off()
  
}, x = lung_transcripts, y = names(lung_transcripts))
```

